---
title: <font color=SteelBlue>Overview</font>
output: 
  html_document:
    code_folding: hide
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
# Load necessary packages. 
library(flexdashboard) # dashboard
library(highcharter) # interactive data visualizations
library(plotly) # interactive data visualizations
library(tidyverse) # metapackge
library(viridis) # color gradients
library(countrycode) # convert country names/codes
library(rjson) # JSON reader
library(crosstalk) # provide interactivity for HTML widgets
library(DT) # display data tables
```


```{r include=FALSE}
# read in data
data = read_csv('./data/who_suicide_statistics.csv') %>%
  filter(country != 'Dominica',
         country != 'Saint Kitts and Nevis')
data = data[complete.cases(data), ]

         
# fix the names of some of the countries in our data to match the country names used by our map later on so that they'll be interpreted and displayed 
data = data %>%
  mutate(country = fct_recode(country, "The Bahamas" = "Bahamas"),
         country = fct_recode(country, "Cape Verde" = "Cabo Verde"),
         country = fct_recode(country, "South Korea" = "Republic of Korea"),
         country = fct_recode(country, "Russia" = "Russian Federation"),
         country = fct_recode(country, "Republic of Serbia" = "Serbia"),
         country = fct_recode(country, "United States of America" = "United States"))

# reorder levels of age to be in chronological order 
data$age = factor(data$age, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))
```

```{r include=FALSE}
# create a custom theme for the plots 
custom_theme = hc_theme(
  colors = c('#5CACEE', 'green', 'red'),
  chart = list(
         backgroundColor = '#FAFAFA', 
         plotBorderColor = "black"),
  xAxis = list(
         gridLineColor = "E5E5E5", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#E5E5E5", 
         minorGridLineColor = "#E5E5E5", 
         tickColor = "#E5E5E5", 
         title = list(style = list(color = "#333333"))), 
  yAxis = list(
         gridLineColor = "#E5E5E5", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#E5E5E5", 
         minorGridLineColor = "#E5E5E5", 
         tickColor = "#E5E5E5", 
         tickWidth = 1, 
         title = list(style = list(color = "#333333"))),   
  title = list(style = list(color = '#333333', fontFamily = "Lato")),
  subtitle = list(style = list(color = '#666666', fontFamily = "Lato")),
  legend = list(
         itemStyle = list(color = "#333333"), 
         itemHoverStyle = list(color = "#FFF"), 
         itemHiddenStyle = list(color = "#606063")), 
  credits = list(style = list(color = "#666")),
  itemHoverStyle = list(color = 'gray'))
```

### <font color=SteelBlue face="Helvetica">**Why do we care about suicide?**</font>

Last month, the suicide of South Korean singers _Sulli_ and _Goo Hara_ shocked the world. 

<img src = "image/sullihara.jpg" alt = "sullihara">

Beyond shock, we hope to find out what kind of people have high risk of suicide and why do people commit suicide.

### <font color=SteelBlue face="Helvetica">**Is suicide a worldwide public health problem?**</font>


*Key facts* [from WHO](https://www.who.int/en/news-room/fact-sheets/detail/suicide)

* Close to <span style="color:red;">800,000</span> people die due to suicide every year.
* Suicide is the <span style="color:red;">third</span> leading cause of death in 15-19-year-olds.
* <span style="color:red;">79%</span> of global suicides occur in low- and middle-income countries.

### <font color=SteelBlue face="Helvetica">**Who is at risk?**</font>

* People with mental disorders (in particular, depression and alcohol use disorders)
* Many suicides also happen impulsively in moments of crisis with a breakdown in the ability to deal with life stresses, such as financial problems, relationship break-up or chronic pain and illness.
* Experiencing conflict, disaster, violence, abuse, or loss and a sense of isolation are strongly associated with suicidal behaviour.
* Suicide rates are also high amongst vulnerable groups who experience discrimination, such as refugees and migrants; indigenous peoples; lesbian, gay, bisexual, transgender, intersex (LGBTI) persons; and prisoners.

The plot below shows suicides by continent and age from 1979 to 2016.

```{r}
# create tibble for our line plot  
overall_tibble = data %>%
  select(year, suicides_no, population) %>%
  group_by(year) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))
# pick color for graph
age_color = rev(plasma(6))
# create new column in data for continent 
# use countrycode() to extract continents from country names
data$continent = countrycode(sourcevar = data$country,
                             origin = "country.name",
                             destination = "continent") 

# reclassify countries that have been coded as 'Americas' into 'North America' and 'South America' by countrycode()
south_america = c('Argentina', 'Brazil', 'Chile', 'Colombia', 'Ecuador', 'Guyana', 'Paraguay', 'Suriname', 'Uruguay')

data$continent[data$country %in% south_america] = 'South America'
data$continent[data$continent == 'Americas'] = 'North America'

# create a tibble for continent and sex
continent_age_tibble = data %>%
  select(continent, age, suicides_no, population) %>%
  group_by(continent, age) %>%
  summarize(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2)) 

continent_age_tibble = continent_age_tibble[complete.cases(continent_age_tibble), ]

# create histogram of suicides by continent.
highchart() %>%
  hc_add_series(continent_age_tibble, mapping = hcaes(x = continent, y = suicide_capita, group = age), type = "column")  %>% 
  hc_colors(colors = age_color) %>%
  hc_title(text = "Suicides by continent and <b>Age</b>", style = (list(fontSize = '14px'))) %>% 
  hc_subtitle(text = "1979-2016") %>%
  hc_tooltip(borderWidth = 1.5, pointFormat = paste("Age: <b> {point.age} </b> <br> Suicides: <b>{point.y}</b>")) %>%
  hc_xAxis(categories = c("Africa", "Asia", "Europe", "North <br> America", "Oceania", "South <br> America"), labels = list(style = list(fontSize = 8))) %>%
  hc_yAxis(labels = list(style = list(fontSize = 10)),
           title = list(text = "Suicides per 100K people",
                        style = list(fontSize = 10)),
           plotLines = list(
             list(color = "black", width = 1, dashStyle = "Dash", 
                  value = mean(overall_tibble$suicide_capita),
                  label = list(text = "Mean = 13.09", style = list(color = "black", fontSize = 6))))) %>%    
  hc_legend(verticalAlign = 'top', enabled = FALSE) %>% 
  hc_add_theme(custom_theme)

```

### <font color=SteelBlue face="Helvetica">**We should pay more attention to suicide.**</font>

Globally, the availability and quality of data on suicide and suicide attempts is poor. Only 80 Member States have good-quality vital registration data that can be used directly to estimate suicide rates. The system must improve to monitor suicide incidence so that to prevent and control it.


The plots shows the suicide by country.

```{r}
# create a tibble with suicide per capita by country for 1985-2015
country_tibble = data %>%
  select(country, suicides_no, population) %>%
  group_by(country) %>%
  summarize(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))

# create interactive world map
highchart() %>%
  hc_add_series_map(worldgeojson, country_tibble, value = "suicide_capita", joinBy = c('name','country'))  %>% 
  hc_colorAxis(stops = color_stops()) %>% 
  hc_title(text = "Suicides by Country") %>% 
  hc_subtitle(text = "1979-2016") %>%
  hc_tooltip(borderWidth = 1.5, headerFormat = "", valueSuffix = " suicides (per 100K people)") 
```


