---
title: "Model"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # general
library(ggalt) # dumbbell plots
library(countrycode) # continent
library(rworldmap) # quick country-level heat maps
library(gridExtra) # plots
library(broom) # significant trends within countries
library(plyr)
library(dplyr)
library(lubridate)
library(caTools)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(data.table)
library(tidyr)
library(corrgram)       
library(corrplot)
library(formattable)
library(cowplot)
library(ggpubr)
library(plot3D)
library(plotly)
library(mgcv)
```

## Suicide rate chages with year

```{r, message=FALSE, warning=FALSE}
## read data
suicide <- read_csv("./data/master.csv") %>% 
  janitor::clean_names() %>%
  as.data.frame() 
suicide = suicide %>% 
  mutate(
    continent = countrycode(sourcevar = suicide[, "country"],
                              origin = "country.name",
                              destination = "continent"),
                            gdp_per_capita = gdp_per_capita * 0.001,
                            gdp_per_capita = log(gdp_per_capita)
  )

country_year <- suicide %>%
  group_by(country, year, continent) %>%
  dplyr::summarize(suicides = sum(suicides_no), 
            population = sum(population), 
            suicide_per_100k = (suicides / population) * 100000, 
            gdp_per_capita = mean(gdp_per_capita),
            gdp_for_year = mean(gdp_for_year),
            hdi_for_year = mean(hdi_for_year))

country_year_trends <- country_year %>%
  ungroup() %>%
  nest(-country) %>% # format: country, rest of data (in list column)
  mutate(model = map(data, ~ lm(suicide_per_100k ~ year, data = .)), # for each item in 'data', fit a linear model
         tidied = map(model, tidy)) %>% # tidy each of these into dataframe format - call this list 'tidied'
  unnest(tidied)

country_year_sig_trends <- country_year_trends %>%
  filter(term == "year") %>%
  mutate(p.adjusted = p.adjust(p.value, method = "holm")) %>%
  filter(p.adjusted < .05) %>%
  arrange(estimate)

country_year_sig_trends$country <- factor(country_year_sig_trends$country, 
                                          ordered = T, 
                                          levels = country_year_sig_trends$country)
```

Here are some plots about simple linear regression of suicide rate (suicide per 100k) and year in countries with significant trends (p < 0.05).

```{r}
top12_increasing <- tail(country_year_sig_trends$country, 12)

country_year %>%
  filter(country %in% top12_increasing) %>%
  ggplot(aes(x = year, y = suicide_per_100k, col = country)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ country) + 
  theme(legend.position = "none") + 
  labs(title = "12 Steepest Increasing Trends", 
       subtitle = "Of countries with significant trends (p < 0.05)", 
       x = "Year", 
       y = "Suicides per 100k")
```

From the plots above we can see that South Korea shows the most significant increasing trend - an increase in suicide per 100 k of 0.931 people with every unit increase in year, which is similar to Guyana - an increase in suicide per 100k of 0.925 people with every unit increase in year.

Guyana’s suicide rate increased by 365% between 1998 and 1999 (5.3 to 24.8) which seems questionable - maybe it's because they changed classification method of defining suicide.

```{r}

top12_decreasing <- head(country_year_sig_trends$country, 12)

country_year %>%
  filter(country %in% top12_decreasing) %>%
  ggplot(aes(x = year, y = suicide_per_100k, col = country)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ country) + 
  theme(legend.position = "none") + 
  labs(title = "12 Steepest Decreasing Trends", 
       subtitle = "Of countries with significant trends (p < 0.05)", 
       x = "Year", 
       y = "Suicides per 100k")
```

Estonia shows the most significant decreasing trend - an decrease in suicide per 100 k of 1.31 people with every unit increase in year.

The following plot is about suicide rate (Suicides per 100 k) change per year of countries with significant trends (p < 0.05).

```{r}
p1 = ggplot(country_year_sig_trends, aes(x = country, y = estimate, col = estimate)) + 
  geom_point(stat = 'identity', size = 2) +
  geom_hline(yintercept = 0, col = "grey", size = 1) +
  scale_color_gradient(low = "green", high = "red") +
  geom_segment(aes(y = 0, 
                   x = country, 
                   yend = estimate, 
                   xend = country), size = 1) +
  labs(x = "", y = "Change Per Year (Suicides per 100k)") +
  scale_y_continuous(breaks = seq(-2, 2, 0.2), limits = c(-1.5, 1.5)) +
  theme(legend.position = "none") +
  coord_flip() +
  theme(
    axis.text.y = element_text(size = 6)
  )

ggplotly(p1) %>% 
  layout(title = "suicide rate change per year", 
         yaxis = list(title = "countries with P < 0.05"))
```

We can get a conclusion that nearly a half of all countries have suicide rates that are changing linearly as time progresses. Between those countries that have suicide rate changing linearly as time progresses, there are 32 (2/3) countries have a decreasing trend. Overall, the suicide rate decreases as time progresses.

## Relationship between economy and suicide

```{r,warning=FALSE, message=FALSE}
country_gdp = country_year %>% 
  ungroup(continent)
p2 = ggplot(country_gdp, aes(x = gdp_per_capita, y = suicide_per_100k, color = continent)) + 
  geom_point(size = 1) + 
  scale_x_continuous(labels=scales::dollar_format(prefix="$"), breaks = seq(0, 130000, 10000)) + 
  labs(title = "Correlation between GDP and Suicides rate", 
       x = "log(GDP (per capital)*0.001)", 
       y = "Suicides per 100k") +
  facet_grid(~continent) +
  geom_smooth(method = "lm")

ggplotly(p2)
```

In Africa and Europe, suicide rate decreases as GDP per capital increases. But in Americas and Oceania, suicide rate increases as GDP per capital increases. In asia, the association between GDP per capital and suicide rate is weak.

Next step is to analyse these data in all-country-level. Because these data may have some influential points, so it's better to drop data with Cooks Distance larger than 4/n and then analyse the relationship between GDP per capital and suicide rate.

```{r}
model1 <- lm(suicide_per_100k ~ gdp_per_capita, data = country_gdp)

gdp_suicide_no_outliers <- model1 %>%
  augment() %>%
  arrange(desc(.cooksd)) %>%
  filter(.cooksd < 4/nrow(.)) %>% # removes 5/93 countries
  inner_join(country_gdp, by = c("suicide_per_100k", "gdp_per_capita")) %>%
  select(country, continent, gdp_per_capita, suicide_per_100k)

model2 <- lm(suicide_per_100k ~ gdp_per_capita, data = gdp_suicide_no_outliers)

summary(model2) %>% 
  tidy() 
```

After drop influential points, the p-value is smaller than 0.05, so we can say that at 0.05 significance level, GDP per capital has association with its suicide rate in all-country-level.

## Relationship between HDI and suicide

```{r,message=FALSE, warning=FALSE}
country_gdp_no_hdi = country_year %>% 
  drop_na() %>% 
  mutate(
    hdi_for_year = mean(hdi_for_year)
  )
p3 = ggplot(country_gdp_no_hdi, aes(x = hdi_for_year, y = suicide_per_100k, color = continent)) + 
  geom_point(size = 1) + 
  scale_x_continuous(labels=scales::dollar_format(prefix="$"), breaks = seq(0, 130000, 10000)) + 
  labs(title = "Correlation between HDI and Suicides rate", 
       x = "HDI", 
       y = "Suicides per 100k", 
       col = "continent") +
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid(~continent) +
  geom_smooth(method = "lm")

ggplotly(p3)
```

In Europe, suicide rate decreases as HDI increases. But in other continent, suicide rate increases as HDI increases. 

As mentioned above, next step is to drop data with Cooks Distance larger than 4/n and then analyse the relationship between HDI and suicide rate in all-country-level.

```{r}
model3 <- lm(suicide_per_100k ~ hdi_for_year, data = country_gdp_no_hdi)

hdi_suicide_no_outliers <- model3 %>%
  augment() %>%
  arrange(desc(.cooksd)) %>%
  filter(.cooksd < 4/nrow(.)) %>% # removes 5/93 countries
  inner_join(country_gdp_no_hdi, by = c("suicide_per_100k", "hdi_for_year")) %>%
  select(country, continent, hdi_for_year, suicide_per_100k)

model4 <- lm(suicide_per_100k ~ hdi_for_year, data = hdi_suicide_no_outliers)

summary(model4) %>% 
  tidy() 
```

After drop the influential point, the p-value is smaller than 0.05, so we can say that at 0.05 significance level, HDI has association with its suicide rate in all-country-level

## Correlation between variables

```{r, message=FALSE, warning=FALSE}
## clean the data
happiness_2015 <- read_csv("./data/2015.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = 2015
  ) %>% 
  select(-region, -standard_error)
happiness_2016 <- read_csv("./data/2016.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = 2016
  ) %>% 
  select(-region, -lower_confidence_interval, -upper_confidence_interval)
happiness_2017 <- read_csv("./data/2017.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = 2017
  ) %>% 
  select(-whisker_high, -whisker_low) 

happiness_2015 = happiness_2015[, c(1,11,2,3,4,5,6,7,8,9,10)]
happiness_2016 = happiness_2016[, c(1,11,2,3,4,5,6,7,8,9,10)]
happiness_2017 = happiness_2017[, c(1,11,2,3,4,5,6,7,9,8,10)]
happiness = rbind(happiness_2015, happiness_2016, happiness_2017)

```

```{r,message=FALSE, warning=FALSE}
overall_data = merge(happiness, suicide, by = c("country", "year")) %>% 
  group_by(country, year, continent) %>%
  dplyr::summarize(suicides = sum(suicides_no), 
            population = sum(population), 
            suicide_per_100k = (suicides / population) * 100000, 
            gdp_capita = mean(gdp_per_capita),
            hap_score = mean(happiness_score),
            family =  mean(family),
            life_expec = mean(health_life_expectancy),
            freedom = mean(freedom),
            trust = mean(trust_government_corruption),
            generosity = mean(generosity),
            residual = mean(dystopia_residual)
            )

asia_overall = overall_data %>% 
  filter(continent == "Asia") 

europe_overall = overall_data %>% 
  filter(continent == "Europe") 
```

**correlation between suicide rate and other variables in all countries**

```{r,message=FALSE, warning=FALSE}
Num.cols <- sapply(overall_data, is.numeric)
Cor.data <- cor(overall_data[, Num.cols])
corrplot(Cor.data, method = 'square', addCoef.col = "black", tl.col = "black", tl.srt = 45, diag = FALSE, number.cex = .7) 
```

Countries in different continent may have different outcomes, which leading to the uncorrelation between suicide rate with other varaibles. So it's more accurate to analyse these data in continent level.

Given that there are few data about Africa and Oceania (1 data point in Oceania and 3 data point in Africa), so only three continents - Americas, Asia and Europe can be analysed.

**correlation between suicide rate and other variables in Asia**

```{r,message=FALSE, warning=FALSE}
Num.cols_asia <- sapply(asia_overall, is.numeric)
Cor.data_asia <- cor(asia_overall[, Num.cols_asia])
corrplot(Cor.data_asia, method = 'square', addCoef.col = "black", tl.col = "black", tl.srt = 45, diag = FALSE, number.cex = .7)
```

Suicide rate in Asia has a slightly positive correlation with degree of satisfiction with family, freedom, but has a slightly negtive correlation with trust of government and gdp per capital.

**correlation between suicide rate and other variables in Europe**

```{r}
Num.cols_europe <- sapply(europe_overall, is.numeric)
Cor.data_europe <- cor(europe_overall[, Num.cols_europe])
corrplot(Cor.data_europe, method = 'square', addCoef.col = "black", tl.col = "black", tl.srt = 45, diag = FALSE, number.cex = .7)
```

Suicide rate in Europe has a negative correlation with gdp per capital, happiness score, degree of satisfiction with family, freedom, trust of government and generasity, which means that if a country in Europe is more prosperous and has a relatively higher happiness score, it is more likely to have a lower suicide rate.

**correlation between suicide rate and other variables in Americas**

```{r,message=FALSE, warning=FALSE}
americas_overall = overall_data %>% 
  filter(continent == "Americas") 
Num.cols_america <- sapply(americas_overall, is.numeric)
Cor.data_america <- cor(americas_overall[, Num.cols_america])
corrplot(Cor.data_america, method = 'square', addCoef.col = "black", tl.col = "black", tl.srt = 45, diag = FALSE, number.cex = .7)

```

Suicide rate in America has a positive correlation with gdp per capital, happiness score, degree of satisfiction with family, freedom, trust of government and generasity, which means that if a country in Americas is more prosperous and has a relatively higher happiness score, it is more likely to have a higher suicide rate.


