---
title: "Model"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # general
library(ggalt) # dumbbell plots
library(countrycode) # continent
library(rworldmap) # quick country-level heat maps
library(gridExtra) # plots
library(broom) # significant trends within countries
library(plyr)
library(dplyr)
library(lubridate)
library(caTools)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(data.table)
library(tidyr)
library(corrgram)       
library(corrplot)
library(formattable)
library(cowplot)
library(ggpubr)
library(plot3D)
library(plotly)
library(mgcv)
```

## Suicide rate chages with year within every country

```{r, message=FALSE, warning=FALSE}
## read data
suicide <- read_csv("./data/master.csv") %>% 
  janitor::clean_names() %>%
  as.data.frame() 
suicide = suicide %>% 
  mutate(
    continent = countrycode(sourcevar = suicide[, "country"],
                              origin = "country.name",
                              destination = "continent"),
                            gdp_per_capita = gdp_per_capita * 0.001,
                            gdp_per_capita = log(gdp_per_capita)
  )

country_year <- suicide %>%
  group_by(country, year, continent) %>%
  dplyr::summarize(suicides = sum(suicides_no), 
            population = sum(population), 
            suicide_per_100k = (suicides / population) * 100000, 
            gdp_per_capita = mean(gdp_per_capita),
            gdp_for_year = mean(gdp_for_year),
            hdi_for_year = mean(hdi_for_year))

country_year_trends <- country_year %>%
  ungroup() %>%
  nest(-country) %>% # format: country, rest of data (in list column)
  mutate(model = map(data, ~ lm(suicide_per_100k ~ year, data = .)), # for each item in 'data', fit a linear model
         tidied = map(model, tidy)) %>% # tidy each of these into dataframe format - call this list 'tidied'
  unnest(tidied)

country_year_sig_trends <- country_year_trends %>%
  filter(term == "year") %>%
  mutate(p.adjusted = p.adjust(p.value, method = "holm")) %>%
  filter(p.adjusted < .05) %>%
  arrange(estimate)

country_year_sig_trends$country <- factor(country_year_sig_trends$country, 
                                          ordered = T, 
                                          levels = country_year_sig_trends$country)
```

Here are some plots about simple linear regression of suicide rate (suicide per 100k) with year in countries with significant trends (p < 0.05).

```{r}
top12_increasing <- tail(country_year_sig_trends$country, 12)

country_year %>%
  filter(country %in% top12_increasing) %>%
  ggplot(aes(x = year, y = suicide_per_100k, col = country)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ country) + 
  theme(legend.position = "none") + 
  labs(title = "12 Steepest Increasing Trends", 
       subtitle = "Of countries with significant trends (p < 0.05)", 
       x = "Year", 
       y = "Suicides per 100k")
```

From the plots above we can see that South Korea shows the most significant increasing trend - an increase in suicide per 100k of 0.931 people with every unit increase in year, which is similar to Guyana - an increase in suicide per 100k of 0.925 people with every unit increase in year.

Guyana’s suicide rate increased by 365% between 1998 and 1999 (5.3 to 24.8) which seems questionable - maybe it's because they changed classification method of defining suicide.)

```{r}

top12_decreasing <- head(country_year_sig_trends$country, 12)

country_year %>%
  filter(country %in% top12_decreasing) %>%
  ggplot(aes(x = year, y = suicide_per_100k, col = country)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ country) + 
  theme(legend.position = "none") + 
  labs(title = "12 Steepest Decreasing Trends", 
       subtitle = "Of countries with significant trends (p < 0.05)", 
       x = "Year", 
       y = "Suicides per 100k")
```

Estonia shows the most significant decreasing trend - an decrease in suicide per 100k of 1.31 people with every unit increase in year.

The following plot is about suicide rate (Suicides per 100k) change per year of countries with significant trends (p < 0.05).

```{r}
p1 = ggplot(country_year_sig_trends, aes(x = country, y = estimate, col = estimate)) + 
  geom_point(stat = 'identity', size = 4) +
  geom_hline(yintercept = 0, col = "grey", size = 1) +
  scale_color_gradient(low = "green", high = "red") +
  geom_segment(aes(y = 0, 
                   x = country, 
                   yend = estimate, 
                   xend = country), size = 1) +
  labs(x = "", y = "Change Per Year (Suicides per 100k)") +
  scale_y_continuous(breaks = seq(-2, 2, 0.2), limits = c(-1.5, 1.5)) +
  theme(legend.position = "none") +
  coord_flip() +
  theme(
    axis.text.y = element_text(size = 4)
  )

ggplotly(p1) %>% 
  layout(title = "suicide rate change per year", 
         yaxis = list(title = "countries with P < 0.05"))
```

We can get a conclusion that nearly a half of all countries have suicide rates that are changing linearly as time progresses. Between those countries that have suicide rate changing linearly as time progresses, there are 32 (2/3) countries have a decreasing trend. Overall, the suicide rate decreases as time progresses.

## Relationship with GDP per capital and suicide rate.

```{r,warning=FALSE, message=FALSE}
country_gdp = country_year %>% 
  ungroup(continent)
p2 = ggplot(country_gdp, aes(x = gdp_per_capita, y = suicide_per_100k)) + 
  geom_point(size = 1, color = "lightblue") + 
  geom_smooth(method = "lm") +
  scale_x_continuous(labels=scales::dollar_format(prefix="$"), breaks = seq(0, 130000, 10000)) + 
  labs(title = "Correlation between GDP and Suicides rate", 
       x = "log(GDP (per capital)*0.001)", 
       y = "Suicides per 100k") +
theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplotly(p2)
```



## Relationship with GDP per capital and suicide rate.

```{r}
country_gdp_no_hdi = country_year %>% 
  drop_na() %>% 
  mutate(
    hdi_for_year = mean(hdi_for_year)
  )
p3 = ggplot(country_gdp_no_hdi, aes(x = hdi_for_year, y = suicide_per_100k)) + 
  geom_point(size = 1, color = "lightblue") + 
  geom_smooth(method = "lm") +
  scale_x_continuous(labels=scales::dollar_format(prefix="$"), breaks = seq(0, 130000, 10000)) + 
  labs(title = "Correlation between HDI and Suicides rate", 
       x = "HDI", 
       y = "Suicides per 100k", 
       col = "continent") +
theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplotly(p3)
```

Happiness score are related to ?

```{r, message=FALSE, warning=FALSE}
## clean the data
happiness_2015 <- read_csv("./data/2015.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = 2015
  ) %>% 
  select(-region, -standard_error) %>% 
  rename(hap_rank = happiness_rank,
         hap_score = happiness_score,
         gdp_capita = economy_gdp_per_capita,
         life_expec = health_life_expectancy,
         trust = trust_government_corruption,
         residual = dystopia_residual
         )
happiness_2016 <- read_csv("./data/2016.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = 2016
  ) %>% 
  select(-region, -lower_confidence_interval, -upper_confidence_interval) %>% 
  rename(hap_rank = happiness_rank,
         hap_score = happiness_score,
         gdp_capita = economy_gdp_per_capita,
         life_expec = health_life_expectancy,
         trust = trust_government_corruption,
         residual = dystopia_residual
         )
happiness_2017 <- read_csv("./data/2017.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    year = 2017
  ) %>% 
  select(-whisker_high, -whisker_low) %>% 
  rename(hap_rank = happiness_rank,
         hap_score = happiness_score,
         gdp_capita = economy_gdp_per_capita,
         life_expec = health_life_expectancy,
         trust = trust_government_corruption,
         residual = dystopia_residual
         )

happiness_2015 = happiness_2015[, c(1,11,2,3,4,5,6,7,8,9,10)]
happiness_2016 = happiness_2016[, c(1,11,2,3,4,5,6,7,8,9,10)]
happiness_2017 = happiness_2017[, c(1,11,2,3,4,5,6,7,9,8,10)]
happiness = rbind(happiness_2015, happiness_2016, happiness_2017)

```

```{r}
overall_data = merge(happiness, suicide, by = c("country", "year")) %>% 
  group_by(country, year, continent) %>%
  dplyr::summarize(suicides = sum(suicides_no), 
            population = sum(population), 
            suicide_per_100k = (suicides / population) * 100000, 
            gdp_capita = mean(gdp_per_capita),
            hap_score = mean(hap_score),
            family =  mean(family),
            life_expec = mean(life_expec),
            freedom = mean(freedom),
            trust = mean(trust),
            generosity = mean(generosity),
            residual = mean(residual)
            )

Num.cols <- sapply(overall_data, is.numeric)
Cor.data <- cor(overall_data[, Num.cols])
corrplot(Cor.data, method = 'square', addCoef.col = "black", tl.col ="black", tl.srt = 45, diag = FALSE, number.cex = .7) 
```




