---
title: "Shiny Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # general
library(ggalt) # dumbbell plots
library(countrycode) # continent
library(rworldmap) # quick country-level heat maps
library(gridExtra) # plots
library(broom) # significant trends within countries
theme_set(theme_light())
library(plyr)
library(dplyr)
library(lubridate)
library(caTools)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(data.table)
library(tidyr)
library(corrgram)       
library(corrplot)
library(formattable)
library(cowplot)
library(ggpubr)
library(plot3D)
library(plotly)
```

```{r}
```{r, message=FALSE, warning=FALSE}
suicide <- read_csv("./data/master.csv") %>% 
  janitor::clean_names() %>%
  as.data.frame() 

suicide = suicide %>% 
  mutate(
    continent = countrycode(sourcevar = suicide[, "country"],
                              origin = "country.name",
                              destination = "continent")
  )


country_year <- suicide %>%
  group_by(country, year, continent) %>%
  dplyr::summarize(suicides = sum(suicides_no), 
            population = sum(population), 
            suicide_per_100k = (suicides / population) * 100000, 
            gdp_per_capita = mean(gdp_per_capita),
            gdp_for_year = mean(gdp_for_year))

```

Column {.sidebar}
-----------------------------------------------------------------------

```{r}

```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}

```

### Chart C

```{r}

```

